Carbonate chemistry at Ny-Ålesund

### 2021-12-23 awipev-CO2.Rmd
pH is OK, normalized too, working on Philipp's Pangaea files

---
title: "Carbonate chemistry at Ny-Ålesund" 
author: "Jean-Pierre Gattuso and Samir Alliouane and Bernard Gentili"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document: default
  word_document: default
  pdf_document: default
  odt_document: default
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

@@

library("RColorBrewer")
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("seacarb")) install.packages("seacarb")
library("seacarb")
if (!require("gridExtra")) install.packages("gridExtra")
library("gridExtra")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("grid")) install.packages("grid")
library(grid)
if (!require("viridis")) install.packages("viridis")
library("viridis")
if (!require("dygraphs")) install.packages("dygraphs")
library("dygraphs")
if (!require("knitr")) install.packages("knitr")
require("knitr")
if (!require("lmodel2")) install.packages("lmodel2")
library("lmodel2")
if (!require("captioner")) install.packages("captioner")
library("captioner")
if (!require("xts")) install.packages("xts")
library("xts")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("skimr")) install.packages("skimr")
library("skimr")
if (!require("FNN")) install.packages("FNN")
library("FNN")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("pander")) install.packages("pander")
library("pander")
if (!require("tsibble")) install.packages("tsibble")
library("tsibble")
if (!require("slider")) install.packages("slider")
library("slider")
if (!require("kableExtra")) install.packages("kableExtra")
library("kableExtra")
if (!require("readxl")) install.packages("readxl")
library("readxl")
if (!require("oce")) install.packages("oce")
library(oce)

@@

@@

@@

kable(quantile(x = tmp$range_diff, na.rm = TRUE), digits = 2, col.names = "Sal. unit", caption = "Quantiles salinity gradient between 0-2 m and > 8 m")
rm(sal_xts)
```

@@

# make the mean of E0 by period of deployement
calib_spectro_all <- group_by(calib_spectro_all, deployment_sf) %>%
  dplyr::summarise(E0int25 = mean(E0int25, na.rm = TRUE),
            E0ext25 = mean(E0ext25, na.rm = TRUE) 
            ) %>% 
  dplyr::filter(!is.na(deployment_sf))
names(calib_spectro_all) <- c("deployment_sf", "E0int25_sf", "E0ext25_sf")

@@

```{r seaFET pH calculation sf_calc all, echo=FALSE, message= FALSE, warning=FALSE}
#### calculation of the calibrated pH with all data irrespective of the depth of the REMOS ####
seafet_calib_all <- sf_calc(calEint = d_all$voltINT, calEext = d_all$voltEXT, E0int25 = d_all$E0int25_sf,  E0ext25 = d_all$E0ext25_sf, calT = d_all$temp_sf, calSal = d_all$sal_insitu_ctd_interp)
# bind new calibrated pH data to the seafet data set.
names(seafet_calib_all) <- c("pHint_tot_sf_all", "pHext_tot_sf_all")
# Despike
d_all <- d_all %>%   
  dplyr::mutate(pHint_tot_sf_all = oce::despike(seafet_calib_all$pHint_tot_sf_all, reference= "median", n=1, k=5001, replace="NA"),
                pHext_tot_sf_all = oce::despike(seafet_calib_all$pHext_tot_sf_all, reference= "median", n=1, k=5001, replace="NA")
  )
# split by depth
d_all <- d_all %>%
  dplyr::mutate(pHint_tot_sf_all_1m = ifelse(depth_group == '0-2', pHint_tot_sf_all, NA),
                pHint_tot_sf_all_3m = ifelse(depth_group == '2-4', pHint_tot_sf_all, NA),
                pHint_tot_sf_all_5m = ifelse(depth_group == '4-6', pHint_tot_sf_all, NA),
                pHint_tot_sf_all_7m = ifelse(depth_group == '6-8', pHint_tot_sf_all, NA),
                pHint_tot_sf_all_10m = ifelse(depth_group == '>8',  pHint_tot_sf_all, NA),
                # depth_discrete = case_when(
                #   !is.na(pHint_tot_sf_all_1m) ~ 1,
                #   !is.na(pHint_tot_sf_all_3m) ~ 3,
                #   !is.na(pHint_tot_sf_all_5m) ~ 5,
                #   !is.na(pHint_tot_sf_all_7m) ~ 7,
                #   !is.na(pHint_tot_sf_all_10m) ~ 10
                #   ),
                # depth_discrete = as.factor(depth_discrete)
                )
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all)))
seafet_depth_xts <- dplyr::select(d_all, datetime, ph_s_sf_temp_insi, pHint_tot_sf_all_1m, pHint_tot_sf_all_3m, pHint_tot_sf_all_5m, pHint_tot_sf_all_7m, pHint_tot_sf_all_10m)
seafet_depth_xts <- as.xts(seafet_depth_xts, order.by = d_all$datetime)
dygraph(seafet_depth_xts, group = "depth", main="pH seaFET", ylab="pHT") %>%
dySeries("pHint_tot_sf_all_10m", label = "pHT 10m", color =  RColorBrewer::brewer.pal(3, "Dark2")[1], strokeWidth = 0, pointSize = 1) %>%
dySeries("pHint_tot_sf_all_7m", label = "pHT 7m", color =  RColorBrewer::brewer.pal(3, "Dark2")[2], strokeWidth = 0, pointSize = 1) %>%
dySeries("pHint_tot_sf_all_5m", label = "pHT 5m", color =  RColorBrewer::brewer.pal(3, "Dark2")[3], strokeWidth = 0, pointSize = 1) %>%
dySeries("pHint_tot_sf_all_3m", label = "pHT 3m", color =  RColorBrewer::brewer.pal(3, "Dark2")[4], strokeWidth = 0, pointSize = 1) %>%
dySeries("pHint_tot_sf_all_1m", label = "pHT 1m", color =  RColorBrewer::brewer.pal(3, "Dark2")[5], strokeWidth = 0, pointSize = 1) %>%
  dySeries("ph_s_sf_temp_insi", label = "pH spectro", color =  "red", strokeWidth = 0, pointSize = 5 ) %>%

@@

seafet_depth_xts <- dplyr::select(d_all, datetime, ph_s_sf_temp_insi, pHint_tot_sf_10m)
seafet_depth_xts <- as.xts(seafet_depth_xts, order.by = d_all$datetime)
dygraph(seafet_depth_xts, group = "depth", main="Spec pH and calibrated seaFET pH at 10 m", ylab="pHT") %>%
dySeries("pHint_tot_sf_10m", label = "pH calib 10m", color =  "black", strokeWidth = 0, pointSize = 1) %>%
dySeries("ph_s_sf_temp_insi", label = "pH spectro", color =  "red", strokeWidth = 0, pointSize = 5 ) %>%
dyEvent(as.POSIXct("2017-08-24 12:00:00", tz="UTC"), "seaFET #1005 installed", labelLoc = "bottom", color="red") %>%
  dyEvent(as.POSIXct("2018-02-02 09:00:00", tz="UTC"), "seaFET settings changed to every 1H", labelLoc = "bottom", color="red") %>% 
  dyEvent(as.POSIXct("2018-04-17 12:00:00", tz="UTC"), "seaFET #0007 installed + continuous mode set", labelLoc = "bottom", color="red") %>%
    dyEvent(as.POSIXct("2019-09-02 12:00:00", tz="UTC"), "seaFET #0007 out of water for cleaning", labelLoc = "bottom", color="red") %>%
  dyShading(from =  sfinterv[1], to= sfinterv[2], color = "#CCCCCC") %>%
  dyShading(from =  sfinterv[2], to= sfinterv[3], color = "#E0E0E0") %>%
  dyShading(from = sfinterv[3], to= Sys.Date(), color = "#CCCCCC") %>%
  dyLegend(show = "auto")   %>%
  dyAxis("y2",valueRange = c(-3, 9)) %>%
  dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
  dyOptions(drawGrid = TRUE, drawPoints = TRUE,useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
  dyRangeSelector(height = 30, dateWindow= c("2017-08-10 00:00:00","2021-01-20 23:00:00"))
rm(seafet_depth_xts)
```

@@

-   Offset between calibrated seaFET data and pH spectro. The offset
    must fit between -0.2/0.2 pH units.

"differences between calibrated sensor pH values and values obtained
from discrete, check samples can be expected to be as small as 0.02 pH
units, and perhaps less." (McLaughlin et al. 2017).

@@

#- pH INT - pH EXT was plotted here.

@@

From now on, one uses seaFET pH from the internal electrode calibrated
with spec pH and seaFET data when the REMOS was below 8 m
(pHint_tot_sf_10m). This unique pH variable is called pHint_tot_sf. All
other pH variables are cleaned up

@@

@@

### **seaFET pH profiles**

```{r seafet profiles, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}

# we look at all depths
tmp <- d_all %>% 
  dplyr::select(year, montht, pHint_tot_sf, depth_group) %>%
  tidyr::drop_na(pHint_tot_sf, depth_group) %>%
  dplyr::group_by(depth_group, year, montht)
summary(aov(tmp$pHint_tot_sf ~ tmp$depth_group * tmp$montht))
p <- tmp %>%
  ggplot(aes(x = depth_group, y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, coef = 1.5) +
  labs(title = "pHT", x = "Depth (m)", y= "pHT seaFET at in situ temperature") +
#  lims(y = c(7.4, 8.3)) +
  coord_cartesian() +
  facet_wrap(vars(montht), scale = "fixed")
print(p)

# we look at 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth) %>%
  dplyr::mutate(depth2 = as.factor(dplyr::case_when(depth < 4 ~ '0-4', 
                                             depth > 8 ~ '8-12'))) %>% 
  dplyr::filter(depth2 == '0-4' | depth2 == '8-12') %>% 
  dplyr::group_by(depth2, montht)
summary(aov(tmp$pHint_tot_sf ~ tmp$depth2 * tmp$montht))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth2), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHT", x = "Depth (m)", y= "pHT seaFET at in situ temperature") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

# we look at 0 to 4 m vs more than 8 m but pH at 4°C

**Conclusion**: The range is significantly larger during some months
when the data deeper than 8 m are considered. Also, pH is sometimes
higher at the surface, at other times it is higher on the bottom (June).
No data are available at 1 m (actually 0 to 2 m). So, to compare surface
and bottom, I combine 1 and 3 m (0 to 4 m) and 9 m (below 8 m).

@@

## Range per months

@@

### **seaFET / durafet pHs (all at 9 m)**

```{r pHs comparaison, echo=FALSE, message= FALSE, warning=FALSE}
# prepare data and remove sf biofouling period
pH_sf <- d_all %>%
  dplyr::select(datetime, sal_insitu_ctd_10m , temp_insitu_ctd_10m, pHint_tot_sf)%>%
  dplyr::filter(!is.na(pHint_tot_sf), !is.na(temp_insitu_ctd_10m), !is.na(sal_insitu_ctd_10m)) %>%
  dplyr::filter(datetime < as.POSIXct("2018-01-01") | datetime > as.POSIXct("2018-02-03"))

pH_sf$pHint_tot_sf_temp_4 <- pHinsi(pH=pH_sf$pHint_tot_sf, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_sf$temp_insitu_ctd_10m, Pinsi=11/10, S=pH_sf$sal_insitu_ctd_10m, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

pH_dur <- d_all %>%
  dplyr::select(datetime, sal_fb,temp_fb, ph_dur_corr)%>%
  dplyr::filter(!is.na(ph_dur_corr), !is.na(temp_fb), !is.na(sal_fb))
pH_dur$pH_durafet_temp_4 <- pHinsi(pH=pH_dur$ph_dur_corr, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_dur$temp_fb, Pinsi=11/10, S=pH_dur$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

# join pH sf norm + pH dur norm
d_all <- left_join(d_all, pH_sf%>%dplyr::select(datetime, pHint_tot_sf_temp_4 ), by="datetime")
d_all <- left_join(d_all, pH_dur%>%dplyr::select(datetime, pH_durafet_temp_4 ), by="datetime")

fit <- lmodel2(data = d_all,  pHint_tot_sf_temp_4 ~ pH_durafet_temp_4 , nperm = 99)
pdur_sf_norm_text <-  ggplot(d_all,aes(x=pH_durafet_temp_4, y= pHint_tot_sf_temp_4, label = monthd))+
  geom_point( aes(color=factor(monthd)), size =1, alpha=0.5) + 
  facet_wrap( ~ factor(monthd), ncol = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  theme(aspect.ratio=1,plot.title = element_text(size=7)) + 
  coord_fixed(ratio = 1 , xlim=c(7.8,8.35), ylim=c(7.8,8.35))+ #scale_colour_discrete(guide="none") +
  labs( x="pH durafet, 4°C", y="pH seaFET, 4°C ", color = "Month ")  +theme_bw()

@@

durafet_sf_contros_xts <- dplyr::select(d_all, datetime, pH_durafet_temp_4, pHint_tot_sf_temp_4)
durafet_sf_contros_xts <- as.xts(durafet_sf_contros_xts, order.by = d_all$datetime)
dygraph(durafet_sf_contros_xts, group = "awipev", main="pH durafet vs pH seaFET at 4°C", ylab="pHT") %>%
    dySeries("pH_durafet_temp_4", label = "pH durafet", color ="blue", strokeWidth=0, pointSize=2) %>%
    dySeries("pHint_tot_sf_temp_4", label = "pH seaFET", color =  "red", strokeWidth = 0, pointSize = 2 ) %>%
    # dyOptions(colors = RColorBrewer::brewer.pal(3, "Dark2"), pointSize = 4) %>%
    dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
       dyLegend(show = "follow")   %>%
    dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = c(1,8,1),useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
    dyRangeSelector(height = 30, dateWindow= NULL) 
rm(durafet_sf_contros_xts, pdur_sf_norm_text)
```

@@

>>>>

---
title: "Carbonate chemistry at Ny-Ålesund" 
author: "Jean-Pierre Gattuso and Samir Alliouane and Bernard Gentili"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  pdf_document: default
  word_document: default
  html_document: default
  odt_document: default
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

@@

library("RColorBrewer")
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("seacarb")) install.packages("seacarb")
library("seacarb")
if (!require("gridExtra")) install.packages("gridExtra")
library("gridExtra")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("grid")) install.packages("grid")
library(grid)
if (!require("viridis")) install.packages("viridis")
library("viridis")
if (!require("dygraphs")) install.packages("dygraphs")
library("dygraphs")
if (!require("knitr")) install.packages("knitr")
require("knitr")
if (!require("lmodel2")) install.packages("lmodel2")
library("lmodel2")
if (!require("captioner")) install.packages("captioner")
library("captioner")
if (!require("xts")) install.packages("xts")
library("xts")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("skimr")) install.packages("skimr")
library("skimr")
if (!require("FNN")) install.packages("FNN")
library("FNN")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("pander")) install.packages("pander")
library("pander")
if (!require("tsibble")) install.packages("tsibble")
library("tsibble")
if (!require("slider")) install.packages("slider")
library("slider")
if (!require("kableExtra")) install.packages("kableExtra")
library("kableExtra")
if (!require("readxl")) install.packages("readxl")
library("readxl")
if (!require("oce")) install.packages("oce")
library(oce)
if (!require("pangaear")) install.packages("pangaear")
library("pangaear")

@@

# # Many issue with sal_ctd_insi
# # read files made available by Philipp with  quality-controlled data
# https://doi.org/10.1594/PANGAEA.929583
# pg_data("10.1594/PANGAEA.929583")

@@

knitr::kable(quantile(x = tmp$range_diff, na.rm = TRUE), digits = 1, col.names = "Salinity (unit)", caption = "Quantiles salinity gradient between 0-2 m and > 8 m", format = "html", table.attr = "style='width:30%;'") %>% kableExtra::kable_classic(full_width = TRUE, position = "center")

@@

rm(sal_xts)
```

@@

# make the mean of E0 by period of deployement
calib_spectro_all <- group_by(calib_spectro_all, deployment_sf) %>%
  dplyr::summarise(E0int25 = mean(E0int25, na.rm = TRUE),
            E0ext25 = mean(E0ext25, na.rm = TRUE) 
            ) %>% 
  dplyr::filter(!is.na(deployment_sf))
names(calib_spectro_all) <- c("deployment_sf", "E0int25_sf_all", "E0ext25_sf_all")

@@

```{r seaFET pH calculation sf_calc all, echo=FALSE, message= FALSE, warning=FALSE}
#### calculation of the calibrated pH with all data irrespective of the depth of the REMOS ####
seafet_calib_all <- sf_calc(calEint = d_all$voltINT, calEext = d_all$voltEXT, E0int25 = d_all$E0int25_sf_all,  E0ext25 = d_all$E0ext25_sf_all, calT = d_all$temp_sf, calSal = d_all$sal_insitu_ctd_interp)
# bind new calibrated pH data to the seafet data set.
names(seafet_calib_all) <- c("pHint_tot_sf_all", "pHext_tot_sf_all")
# Despike
d_all <- d_all %>%   
  dplyr::mutate(pHint_tot_sf_all = oce::despike(seafet_calib_all$pHint_tot_sf_all, reference= "median", n=1, k=5001, replace="NA"),
                pHext_tot_sf_all = oce::despike(seafet_calib_all$pHext_tot_sf_all, reference= "median", n=1, k=5001, replace="NA")
  )
# split by depth
d_all <- d_all %>%
  dplyr::mutate(pHint_tot_sf_all_1m = ifelse(depth_group == '0-2', pHint_tot_sf_all, NA),
                pHint_tot_sf_all_3m = ifelse(depth_group == '2-4', pHint_tot_sf_all, NA),
                pHint_tot_sf_all_5m = ifelse(depth_group == '4-6', pHint_tot_sf_all, NA),
                pHint_tot_sf_all_7m = ifelse(depth_group == '6-8', pHint_tot_sf_all, NA),
                pHint_tot_sf_all_10m = ifelse(depth_group == '>8',  pHint_tot_sf_all, NA),
                # depth_discrete = case_when(
                #   !is.na(pHint_tot_sf_all_1m) ~ 1,
                #   !is.na(pHint_tot_sf_all_3m) ~ 3,
                #   !is.na(pHint_tot_sf_all_5m) ~ 5,
                #   !is.na(pHint_tot_sf_all_7m) ~ 7,
                #   !is.na(pHint_tot_sf_all_10m) ~ 10
                #   ),
                # depth_discrete = as.factor(depth_discrete)
                )
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all)))
seafet_depth_xts <- dplyr::select(d_all, datetime, ph_s_sf_temp_insi, pHint_tot_sf_all_1m, pHint_tot_sf_all_3m, pHint_tot_sf_all_5m, pHint_tot_sf_all_7m, pHint_tot_sf_all_10m)
seafet_depth_xts <- as.xts(seafet_depth_xts, order.by = d_all$datetime)
dygraph(seafet_depth_xts, group = "depth", main="Spec pH and seaFET pH at all depths (calibrated using all data, irrespective of the depth of the seafet)", ylab="pHT") %>%
dySeries("pHint_tot_sf_all_10m", label = "pHT 10m", color =  RColorBrewer::brewer.pal(3, "Dark2")[1], strokeWidth = 0, pointSize = 1) %>%
dySeries("pHint_tot_sf_all_7m", label = "pHT 7m", color =  RColorBrewer::brewer.pal(3, "Dark2")[2], strokeWidth = 0, pointSize = 1) %>%
dySeries("pHint_tot_sf_all_5m", label = "pHT 5m", color =  RColorBrewer::brewer.pal(3, "Dark2")[3], strokeWidth = 0, pointSize = 1) %>%
dySeries("pHint_tot_sf_all_3m", label = "pHT 3m", color =  RColorBrewer::brewer.pal(3, "Dark2")[4], strokeWidth = 0, pointSize = 1) %>%
dySeries("pHint_tot_sf_all_1m", label = "pHT 1m", color =  RColorBrewer::brewer.pal(3, "Dark2")[5], strokeWidth = 0, pointSize = 1) %>%
  dySeries("ph_s_sf_temp_insi", label = "pH spectro", color =  "red", strokeWidth = 0, pointSize = 5 ) %>%

@@

seafet_depth_xts <- dplyr::select(d_all, datetime, ph_s_sf_temp_insi, pHint_tot_sf_10m)
seafet_depth_xts <- as.xts(seafet_depth_xts, order.by = d_all$datetime)
dygraph(seafet_depth_xts, group = "depth", main="Spec pH and seaFET pH at 10 m (calibrated with 10 mdata only)", ylab="pHT") %>%
dySeries("pHint_tot_sf_10m", label = "pH calib 10m", color =  "black", strokeWidth = 0, pointSize = 1) %>%
dySeries("ph_s_sf_temp_insi", label = "pH spectro", color =  "red", strokeWidth = 0, pointSize = 5 ) %>%
dyEvent(as.POSIXct("2017-08-24 12:00:00", tz="UTC"), "seaFET #1005 installed", labelLoc = "bottom", color="red") %>%
  dyEvent(as.POSIXct("2018-02-02 09:00:00", tz="UTC"), "seaFET settings changed to every 1H", labelLoc = "bottom", color="red") %>% 
  dyEvent(as.POSIXct("2018-04-17 12:00:00", tz="UTC"), "seaFET #0007 installed + continuous mode set", labelLoc = "bottom", color="red") %>%
    dyEvent(as.POSIXct("2019-09-02 12:00:00", tz="UTC"), "seaFET #0007 out of water for cleaning", labelLoc = "bottom", color="red") %>%
  dyShading(from =  sfinterv[1], to= sfinterv[2], color = "#CCCCCC") %>%
  dyShading(from =  sfinterv[2], to= sfinterv[3], color = "#E0E0E0") %>%
  dyShading(from = sfinterv[3], to= Sys.Date(), color = "#CCCCCC") %>%
  dyLegend(show = "auto")   %>%
  dyAxis("y2",valueRange = c(-3, 9)) %>%
  dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
  dyOptions(drawGrid = TRUE, drawPoints = TRUE,useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
  dyRangeSelector(height = 30, dateWindow= c("2017-08-10 00:00:00","2021-01-20 23:00:00"))
rm(seafet_depth_xts)
```

@@

Offset between calibrated seaFET data and pH spectro. The offset must be between -0.2 and 0.2 pH units. "differences between calibrated sensor pH values and values obtained from discrete, check samples can be expected to be as small as 0.02 pH units, and perhaps less." (McLaughlin et al. 2017).

@@

@@

pH INT - pH EXT was plotted here. From now on, one uses seaFET pH from the internal electrode calibrated
with spec pH and seaFET data when the REMOS was below 8 m (pHint_tot_sf_10m). This unique pH variable is called pHint_tot_sf. All other pH variables are cleaned up. We also calculate pHint_tot_sf. at 4°C: pHint_tot_sf_temp_4.

@@

# normalized seafet pH 4°C and remove sf biofouling period
tmp <- d_all %>%
  dplyr::select(datetime, depth, at, sal_insitu_ctd, temp_sf, pHint_tot_sf) %>%
  tidyr::drop_na() %>%
  dplyr::filter(datetime < as.POSIXct("2018-01-01") | datetime > as.POSIXct("2018-02-03"))
tmp$pHint_tot_sf_temp_4 <- seacarb::pHinsi(pH=tmp$pHint_tot_sf, ALK=mean(tmp$at*1e-6, na.rm=TRUE), Tinsi=4, Tlab=tmp$temp_sf, Pinsi=tmp$depth/10, S=tmp$sal_insitu_ctd, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")
d_all <- d_all %>% 
  left_join(tmp %>% dplyr::select(datetime, pHint_tot_sf_temp_4), by="datetime")
```

<!-- ### **seaFET pH profiles** -->

```{r seafet profiles, eval=FALSE, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}

# # we look at all depths
# tmp <- d_all %>% 
#   dplyr::select(year, montht, pHint_tot_sf, depth_group) %>%
#   tidyr::drop_na(pHint_tot_sf, depth_group) %>%
#   dplyr::group_by(depth_group, year, montht)
# summary(aov(tmp$pHint_tot_sf ~ tmp$depth_group * tmp$montht))
# p <- tmp %>%
#   ggplot(aes(x = depth_group, y = pHint_tot_sf)) +
#   geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, coef = 1.5) +
#   labs(title = "pHT", x = "Depth (m)", y= "pHT seaFET at in situ temperature") +
# #  lims(y = c(7.4, 8.3)) +
#   coord_cartesian() +
#   facet_wrap(vars(montht), scale = "fixed")
# print(p)
# 
# # we look at 0 to 4 m vs more than 8 m
# tmp <- d_all %>% 
#   dplyr::select(montht, pHint_tot_sf, depth) %>%
#   dplyr::mutate(depth2 = as.factor(dplyr::case_when(depth < 4 ~ '0-4', 
#                                              depth > 8 ~ '8-12'))) %>% 
#   dplyr::filter(depth2 == '0-4' | depth2 == '8-12') %>% 
#   dplyr::group_by(depth2, montht)
# summary(aov(tmp$pHint_tot_sf ~ tmp$depth2 * tmp$montht))
# 
# p <- tmp %>% 
#   ggplot(aes(x=as.factor(depth2), y = pHint_tot_sf)) +
#   geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
#   labs(title = "pHT", x = "Depth (m)", y= "pHT seaFET at in situ temperature") +
#   facet_wrap(vars(montht), scale = "free_y")
# print(p)
# 
# # we look at 0 to 4 m vs more than 8 m but pH at 4°C

@@

<!-- **Conclusion**: The range is significantly larger during some months -->
<!-- when the data deeper than 8 m are considered. Also, pH is sometimes -->
<!-- higher at the surface, at other times it is higher on the bottom (June). -->
<!-- No data are available at 1 m (actually 0 to 2 m). So, to compare surface -->
<!-- and bottom, I combine 1 and 3 m (0 to 4 m) and 9 m (below 8 m). -->

<!-- Depth and month of the year have a statistically significant effect on -->
<!-- pH. The interaction is also significant, indicating that the effect of -->
<!-- depth differs depending on the month. -->

<!-- -   To compare with pCO2/pCO2 measured in the Ferrybox seems therefore -->
<!--     better to stick with pH_10m, that is at depths deeper than 8 m. -->
<!-- -   pCO2 would have the same behaviour as pH. For air-sea CO2 fluxes, -->
<!--     one should use data from 0 to 4 m. -->

@@

```{r durafet clean up, echo=FALSE, message= FALSE, warning=FALSE}
d_all <- d_all %>% 
  dplyr::mutate(pH_dur = ph_dur_corr) %>% # the right durafet pH at in situ temp from now on
  dplyr::select(-ph_dur, -ph_dur_corr)
# normalized durafet pH 4°C
tmp <- d_all %>%
  dplyr::select(datetime, at, sal_fb, temp_fb, pH_dur) %>%
  tidyr::drop_na()
tmp$pH_dur_temp_4 <- seacarb::pHinsi(pH=tmp$pH_dur, ALK=mean(tmp$at*1e-6, na.rm=TRUE), Tinsi=4, Tlab=tmp$temp_fb, Pinsi=11/10, S=tmp$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")
d_all <- d_all %>% 
  left_join(tmp %>% dplyr::select(datetime, pH_dur_temp_4 ), by="datetime")
```
## Range per months

@@

### **seaFET / durafet pHs (all > 8 m)**

```{r pHs comparaison 4°C, echo=FALSE, message= FALSE, warning=FALSE}
# prepare data and remove sf biofouling period
pH_sf <- d_all %>%
  dplyr::select(datetime, sal_insitu_ctd_10m, temp_insitu_ctd_10m, pHint_tot_sf_temp_4, depth_group) %>%
  #tidyr::drop_na() %>%
  dplyr::filter(datetime < as.POSIXct("2018-01-01") | datetime > as.POSIXct("2018-02-03")) %>% 
  dplyr::mutate(pHint_tot_sf_temp_4_10m = ifelse(depth_group == ">8", pHint_tot_sf_temp_4, NA))

#pH_sf$pHint_tot_sf_temp_4 <- pHinsi(pH=pH_sf$pHint_tot_sf, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_sf$temp_insitu_ctd_10m, Pinsi=11/10, S=pH_sf$sal_insitu_ctd_10m, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

pH_dur <- d_all %>%
  dplyr::select(datetime, sal_fb,temp_fb, ph_dur_corr)%>%
  tidyr::drop_na() %>%
pH_dur$pH_dur_temp_4 <- pHinsi(pH=pH_dur$ph_dur_corr, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_dur$temp_fb, Pinsi=11/10, S=pH_dur$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

# add pH sf and pH dur normalized at 4°C ti d_all
#d_all <- left_join(d_all, pH_sf%>%dplyr::select(datetime, pHint_tot_sf_temp_4 ), by="datetime")
d_all <- d_all %>% 
  left_join(pH_dur %> %dplyr::select(datetime, pH_dur_temp_4), by="datetime")

fit <- lmodel2(data = d_all,  pHint_tot_sf_temp_4 ~ pH_dur_temp_4, nperm = 99)
pdur_sf_norm_text <-  ggplot(d_all,aes(x=pH_dur_temp_4, y= pHint_tot_sf_temp_4, label = monthd))+
  geom_point( aes(color=factor(monthd)), size =1, alpha=0.5) + 
  facet_wrap( ~ factor(monthd), ncol = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  theme(aspect.ratio=1,plot.title = element_text(size=7)) + 
  coord_fixed(ratio = 1 , xlim=c(7.8,8.35), ylim=c(7.8,8.35)) + #scale_colour_discrete(guide="none") +
  labs( x="pH durafet, 4°C", y="pH seaFET, 4°C ", color = "Month ")  +theme_bw()

@@

durafet_sf_contros_xts <- dplyr::select(d_all, datetime, pH_dur_temp_4, pHint_tot_sf_temp_4)
durafet_sf_contros_xts <- as.xts(durafet_sf_contros_xts, order.by = d_all$datetime)
dygraph(durafet_sf_contros_xts, group = "awipev", main="pH durafet vs pH seaFET at 4°C", ylab="pHT") %>%
    dySeries("pH_dur_temp_4", label = "pH durafet", color ="blue", strokeWidth=0, pointSize=2) %>%
    dySeries("pHint_tot_sf_temp_4", label = "pH seaFET", color =  "red", strokeWidth = 0, pointSize = 2 ) %>%
    # dyOptions(colors = RColorBrewer::brewer.pal(3, "Dark2"), pointSize = 4) %>%
    dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
       dyLegend(show = "follow")   %>%
    dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = c(1,8,1),useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
    dyRangeSelector(height = 30, dateWindow= NULL) 
rm(durafet_sf_contros_xts, pdur_sf_norm_text)
```

@@

**Conclusion for pH**: The range is significantly larger during some months
when the data deeper than 8 m are considered. Also, pH is sometimes
higher at the surface, at other times it is higher on the bottom (June).
No data are available at 1 m (actually 0 to 2 m). So, to compare surface
and bottom, I combine 1 and 3 m (0 to 4 m) and 9 m (below 8 m).
